---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

This is tutorial 1. We are working on assignment 0.

```{r}
mtcars
```


#ctrl + alt + i <-windows; generates a new chunk.
#command + option + i <-mac; generates a new chunk



#ctrl + shift + m; generates a new tidyverse pipepine windows
#command + shift  + m; generates a tidyverse pipeline for mac


```{r}
#base R 
summary(mtcars)
```

Let us have a look at the poor man's boxplot within base R. 

```{r}
boxplot(mtcars$mpg)

#attach the dataframe to my global environment 
attach(mtcars)
boxplot(mpg)

#detach allows you to remove dataframe from global environment
```

Let's play with the tidyverse functions.

```{r, message=F, warning=F}
library(tidyverse)
ggplot(mtcars, aes(x=hp, y=mpg))+geom_point(col="blue")+geom_smooth(method="lm", col="red")+xlab("Gross horsepower")+ylab("Miles/(US) gallon")
```
Let's try to go back and create the boxplot for  mpg

```{r}
ggplot(mtcars, aes(x=mpg))+geom_boxplot()+xlab("Miles/(US) gallon")
```
How do we download data frames (external)?

Ans: We have to use readr in tidyverse. Note that read_delim is a function inside of readr. 


```{r}
my_url="http://ritsokiguess.site/datafiles/jumping.txt"
rats=read_delim(my_url," ")
```
Prelude to pipelines: here we are using tidyr.

```{r}
rats%>% group_by(group) %>% summarize(m=mean(density))
```

