---
title: "Tutorial 10"
author: "Nnenna Asidianya"
date: "3/28/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, wwarning = FALSE, message = FALSE)
```


## Categorical variables in multiple regression:

Recall the example from last week we started multiple linear regression in the context of miles per gallon (mpg). 

Suppose we want to examine the association between a continuous response variable (mpg) and a categorical predictor variable (cyl). We learned that a scatter plot is plausible course of action to display the relationship in bivariate statistics.

```{r}
library(tidyverse)
ggplot(mtcars, aes(x=as.factor(cyl), y=mpg))+geom_point()
```


Here we see that cyl is a discrete variable, so the scatter plot display may not necessarily translate as well. Let's change gears. 


```{r}
ggplot(mtcars, aes(x=as.factor(cyl), y=mpg))+geom_boxplot()

model2<-lm(mpg~wt + as.factor(cyl), dat=mtcars)
summary(model2)
```

What is the interpretation of a categorical coefficient?

Here we an see that the reference group is "4" within the cylinders. Why? (We have a decrease in the mpg when you move to 6 and 8 cylinder classes)

Interpretation of the coefficient must be made in comparison to the reference group (4 cylinders), which is absorbed in the intercept coefficient. 

How do we determine significance?

$$H_0: \text{All}  \hspace{0.05in} \beta_i= 0$$
$$H_1: \text{At least one $\beta_i$ is not 0}$$

This is the same as it always was. 

## Functions in R

An R function is created by using the keyword *function*. The basic syntax of an R function definition is as follows:


```{r}
function_name <- function(arg_1,..) {
   Function_body 
}
```

The different parts of a function are:

* Function Name − This is the actual name of the function. It is stored in the R environment as its name. 

* Arguments − This is an placeholder to pass an input value into your function (i.e. $f(x)$). 

* Function Body − The  function contains the statements that determines what the function does. 

### Example 1:

```{r}
#Create a function to print out the values for perfect squares. 
perfect.squares<-function(x){for (x in 1:10){
  if(x%%x == 0)
    b<-x^2
  print (b)
}}
#pass a value to the function
perfect.squares(10)
```


### Example 2:

```{r}
#Create a function with more than one argument
translate<-function(x,y){
  if (x=="NA"){x=0}
  else if (x!="NA"){x=x}
  print(x)
  if (y== -1){y="NA"}
  else if (y!=-1){y=y}
  print(y)
}
#Testing the function with input values 
x<-"NA"
y<-"-1"
x2<-1
y2<-0
translate(x, y)
translate(x2, y2)
```

As an aside, in this example I could have stuck to the general structure of:

<center>
if (test_expression) {
statement1
} 
else {
statement2
}
</center>


More information from my R Package Workshop: 

https://github.com/asidianya/R-Workshop/blob/main/Functions.pdf


# Assignment 10: Matrix Algebra 

```{r}

buses=read.table('buses.txt', header=FALSE)
colnames(buses) <- c("expense","miles", "percentD", "percentF", "receipt")
attach(buses)

pairs(~ expense +  percentD + miles, data = buses)

## Computing the sums for the predictor and response variables 
X_1=miles

X_2=percentD

X_1_s=sum(X_1)

X_2_s=sum(X_2)

X_1_sq=sum(X_1^2)

X_2_sq=sum(X_2^2)

X_12_s=sum(X_1*X_2)

Y_s=sum(expense)

X_1_Y=sum(X_1*expense)

X_2_Y=sum(X_2*expense)

## Create XY matrix
I=rep(1,24)

X=cbind(I, X_1, X_2)

Y=cbind(expense)

XY=t(X)%*%Y

## Create Matrix XX matrix and find the inverse matrix in R

XX=t(X)%*%X
Inverse=solve(XX)

#Exercise: XX%*%Inverse (should be close to I)

## beta
beta=Inverse%*%XY

fit=X%*%beta #(i.e. Yhat=XB )

res=Y-fit

## Covariance of fitted values. 


X_r1=c(1, 46230, 43.67)
X_r2=c(1, 7360, 65.51)
X_r=rbind(X_r1, X_r2)



CV=X_r%*%Inverse%*%t(X_r)

## Multiple Linear Regression
multi.fit = lm(expense~ miles + percentD, data=buses)
summary(multi.fit)


res.aov <- aov(expense~ miles + percentD, data=buses)
summary(res.aov)

AOV<-anova(multi.fit)

```

