---
title: "Tutorial 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

## Tidyverse 

```{r}
library(tidyverse)


#from tidyverse
fish_encounters
```

How about we pivot wider and look our data. 


```{r}
fish_encounters2<-fish_encounters %>%
  pivot_wider(names_from = station, values_from = seen)
fish_encounters2

#fill in the NAs with 0.

fish_encounters3 <- fish_encounters %>% 
  pivot_wider(names_from = station, values_from = seen, values_fill = 0)
fish_encounters3
```

Note: The select() function in tidyverse allows you to select certain columns.  Ex: select(Release| I80_1) select the first two columns; select (fish_encounters[1:2]).In contrast filter() allows you to select certain rows.

Suppose we are interested in looking at fish within the range of 4842 to 4844.

```{r}
fish_encounters3r<-fish_encounters3 %>% filter(fish %in% (4842:4844))
```

Let's go back to the original data set that is in longer format and tidy things up. 

How many fishes are seen within each station?


```{r}
count_fish<-fish_encounters %>%
group_by(station) %>%
summarize(sum_fish = sum(seen))%>%
arrange(desc(sum_fish))

```

Let's go back to our trusty mtcars example!

```{r}
attach(mtcars)
mtcars

#tibble allows us to take columns within a dataframe
#as.data.frame in tidyverse does a similar thing

cylinder<-tibble(cyl, wt, mpg)
cylinder

#for illustrative purposes; this data is messy so we will not use it.
wider<- cylinder %>% pivot_wider(names_from = cyl, values_from = wt) 
```

Suppose we want to use logical operators to subset the data (i.e. ==, !=, <, >, |, &, etc.)

Question: What cyl class is most common for a wt between 2.8 to 3.2?

```{r}
weight<- cylinder %>% filter(wt>=2.8, wt<=3.2)
```


Question: Is there an association between weight (per 1000) of car and miles (US)/gallon?
```{r}
#create a scatterplot for wt and mpg. 
ggplot(cylinder, aes(x=wt, y=mpg))+geom_point()+geom_smooth(method="lm")


```

