---
title: "Tutorial 11"
author: "Nnenna Asidianya"
date: "4/5/2021"
output: html_document
---

#Matrix Algebra in R

Let $A$ and $B$ be two matrices. These are the common operators in R:

* A %*% B is matrix multiplication

* t(A) is transpose

* solve(A)  is the inverse of A, where A is a square matrix. 

* solve(A, b) returns $x  =  A^{-1} y$ where the system is $ y = Ax)

## Example 1: 

```{r}
#This is a 2x3 matrix
 A <- matrix(c(1, 3, 2, 2, 8, 9), ncol = 3)
A
```

```{r}
#This is 3x3 matrix' default is byrow=F so if you want to append the matrix by rows change this

B <- matrix(c(1, 3, 2, 2, 8, 9, 2, 3, 4), ncol = 3)
B
```

We can multiple in one direction but it is not commutative these are not both square matrices (nor inverses).

```{r}

A %*% B

```

## Example 2

Let's consider our matrix $A$ from Example 1:

```{r}
#defines the transpose 3x2 matrix

t(A)
```
This changes the way in which we perform matrix multiplication.

```{r}
B %*% t(A)
```

# Examples 3 and 4

Let's solve the system by introducing a $3 x 1$ vector.

```{r}
y = rbind(2, 4, 6)
y

```


Notice that B is invertible since  I can solve for the inverse:

```{r}
solve(B)
```


Hence we can solve for the system as follows:

```{r}
 x = solve(B, y)
x
```
