---
title: "Tutorial 9"
author: "Nnenna Asidianya"
date: "3/21/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```


Recall from last week that we had the following simple linear regression model using mtcars:

# Simple Linear Regression

```{r}
library(tidyverse)
attach(mtcars)

ggplot(mtcars, aes(x=wt, y=mpg)) +geom_point() +geom_smooth(method="lm")

model<-lm(mpg~wt, mtcars)
summary(model)
```

Note: Two degrees lost to estimate the intercept and slope term, so really the degrees of freedom is $df=n-2$. Therefore:

$P(T>|t|) = 1.29e-10$. This is equivalent to assessing whether $|t|>t_{crit} = t_{0.975; 30}=2.042$

Question: How do we know that the coefficient is significant? Perform hypothesis test:

$$H_0: \beta_1 = 0$$
$$H_1: \beta_1 \neq 0$$. 


Since p<0.05 we reject the null hypothesis and conclude the alternative. There appears to be an association between weight and miles per gallon driven. The more a car weighs the slower the car moves a given mile per gallon.

Interpretation of coefficient: For every unit increase in the weight of a car (1000 lbs) the miles per gallon decreases by 5.34.

## Residual plots

To assess non-constant variance. It appears we have three outliers, and we have some upward U shape. This may indicate some strange non-constant variance somewhere. 

```{r}
plot(model)

#this gives the qqplot and the residual plot
plot(model, 1:2)
```

Let's pull up residuals to make a histogram:

```{r}
#check your outliers positions you can use  the following code (remove the hashtag)
#res2<-model$residuals

res<-residuals(model)

res<-tibble(res)

ggplot(res, aes(x=res)) + geom_histogram(col="black", fill="red", bins = 7)+xlab("residuals")

```


# Multiple Linear Regression

```{r}
ggplot(mtcars, aes(x=as.factor(cyl), y=mpg))+geom_boxplot()

model2<-lm(mpg~wt + as.factor(cyl), dat=mtcars)
summary(model2)
```


What is the interpretation of said model coefficients?

Ans: Holding the cylinder class fixed then the change in the miles per gallon driven as the weight increases by 1000 lbs is  a decease by 3.20. 

How do we determine significance?

$$H_0: \text{All}  \hspace{0.05in} \beta_i= 0$$

$$H_1: \text{At least one}  \hspace{0.05in} \beta_i \neq 0$$. 

Let's assess the residual plots:

```{r}
plot(model2, 1:2)
```

We perhaps had non-constant variance because we were missing information from our simple linear regression model. 
