---
title: "Week 3 R Tutorial"
author: "Nnenna Asidianya"
date: "1/31/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)

attach(mtcars)
#head(mtcars)

#check whether your variable is distributed normally. 

#base R
boxplot(wt)

#tidyverse 
ggplot(mtcars, aes(x=wt))+geom_boxplot()

+coord_flip()

#base R
hist(wt)

#tidyverse
ggplot(mtcars, aes(x=wt))+geom_histogram(fill="red", color="black", bins=15)
```

Question: Is it true that the distribution is normal?

Answer: Noted that the distribution is skewed. In fact, we can see that there is a long left tail, and $Q_3$ is closer to the median than $Q_1$. This may indicate a left skew. We are suspicious of the three upper outliers. 


Now let's check the mean of the weight variable, something we should never do in practice, to "guess" at the population level mean.


I am using the sample mean (i.e. 3.21725) to find a good comparison parameter at the population level. Here I will guess 2. 
```{r}

mean(wt)

# One-sample t-test (note; df=n-1)
#H_0: mu=2, versus mu not equal to 2
alpha=0.05

#percentile for my quantile 
quant = 1-alpha/2
n <-32

#the critical value for t (i.e quantile for td-distribution)
qt(quant, n-1)

res <- t.test(wt, mu = 2)
res

#If we think we have a direction of the effect then how can we code this

#this will change the quantile.

quant_one = 1-alpha 
qt(quant_one, n-1)

res2 <- t.test(wt, mu = 2, alternative= "greater")
res2




```


#two independent groups t-test

```{r}
#independent two-sample t-test
data<-tibble(vs, wt)

group2<-data %>% filter(vs==0) #vshaped engine
group1<-data %>% filter(vs==1) #straight engine



se1<-sd(group1$wt)/sqrt(length(group1$vs)) #(the std.err are comparable)
se2<-sd(group2$wt)/sqrt(length(group2$vs))

#default is two sided Welch test (unequal variances)
t.test(group1$wt, group2$wt)

#two sided pooled t-test
t.test(group1$wt, group2$wt, var.equal = TRUE)

#we can change the nature of the difference by adding mu=

```

I am comparing the two groups vs=0 (V-shaped) and vs=1 (straight). I am comparing them on the dimension of weight. Is there a significant difference in the weight of the two cars?

Question: How do we know if we should do a pooled or a welsh t-test? Note: We do a pooled test if we suspect the two independent groups have an equal variance. Otherwise wehave to use Welch t-test. 

Answer: The standard errors for group 1 and 2 are approximately 0.191 and.213 which are close, so let's try pooled t-test. 

Our p-value is less than 0.05, and t is negative, which indicates that the V shaped engines are significant heavier than straight engines. In the answer if given a two sided test, you should really say "different from" rather than specifying a direction.

Question: Do you think that it was appropriate to use the t-test within the scope of the sample sizes for groups 1 and group 2?

Answer: Check the boxplots. Are they symmetric?
