---
title: "Tutorial 2"
author: "Nnenna Asidianya"
date: "1/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tutorial 2

Please use message=FALSE to hide the loading of the tidyverse package. 

Other neat features may be warning=FALSE. 

In your final report, you may want to use echo=FALSE, if you do not want your code in between your write up. But do not use it for the assignments since we do want to see your code. The default is echo=TRUE. 

```{r, message=FALSE}
library(tidyverse)
attach(mtcars)
#install.packages("skimr")
library(skimr)

view(mtcars) #your first line of defense for any assignment
skim(mtcars)

#as.numeric() transforms numbers that are presented as characters into numerical values. 

#useful if you wish to find our the information about your data frame. It also works for functions you are confused about
help(mtcars)
help(mean)

```
Let's ask ourselves if we can look at the association between weight of a car (wt) and the number of cylinders (cyl).

```{r}
#bivariate analysis - two variables
ggplot(mtcars, aes(wt, cyl)) + geom_point()

#univariate analysis - one variable
ggplot(mtcars, aes(wt)) + geom_histogram()+stat_bin(bins=15)

x<-rnorm(100, mean=1, sd=0.5)
y<-rgamma(100, 1, 1)
z<-runif(100, 0, 1)

hist(x)
hist(y)
hist(z)
```

Question 1: What do notice about your scatterplot? 

Ans: Notice that the scatterplot has three difference cylinder classes from the plot. The weight does appear to be related to the cylinder classes; cars that weight more appear to have more cylinders. To be very specific I can see that there is overlap in the lower level cylinder classes around 3000 pounds.  

Question 2: What is the shape of the distribution for weight (wt)? Plot the histogram using ggplot. 

Ans: The data looks somewhat uniform based on the difficulty of deciphering the mode; the frequency looks somewhat constant across each bin.

## Extra example: Assignment 0

```{r}
#lm stands for linear model. 

#y ~ x -> y given x (i.e. y=mx+b)

ggplot(mtcars, aes(x=hp, y=mpg))+geom_point(col="red")+ geom_smooth(method="lm")

#curiousity is a good thing!

fit<-lm(mpg~hp, mtcars)
summary(fit)

#checking for outliers
ggplot(mtcars, aes(x=hp)) + geom_boxplot()

ggplot(mtcars, aes(x=hp))+ geom_histogram()+stat_bin(bins=15)
```

Question 3 : What do you notice about Miles (mpg) as a function of horsepower (hp)?

Ans: I notice a negative (1) linear (?) (2) relationship.  The effect looks at least moderate (3) between the variables. High horsepower leads to lower miles/gallon (hp) (4).

**Include all these four things to get full marks**

*TA's homework: Can I write this line on ggplot? Yes, I will have a look for next week. *

Question 4: Make a boxplot of hp. What do you notice about the shape? Describe any other important features?

Ans: I notice an outlier at greater than 300 horsepower. Notice that the distribution is right skewed since the median (black line) is closer to Q1 than Q3 and we have a long right tail. 