---
title: "Tutorial 10"
author: "Nnenna Asidianya"
date: "3/28/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, wwarning = FALSE, message = FALSE)
```

##Multiple Linear regression

a) Read in the data and display it (or at least the first ten lines).

```{r}
library(tidyverse)
gpa<-read_table2("http://ritsokiguess.site/datafiles/gpa.txt")
attach(gpa)
```


b)  Make a scatterplot of high school GPA against university GPA. Which variable should be the response and which explanatory? Explain briefly. Add a smooth trend to your plot.

```{r}
library(gridExtra)

p1<-ggplot(gpa, aes(high_GPA, univ_GPA))+geom_point()+geom_smooth(method="lm")

p2<-ggplot(gpa, aes(math_SAT, univ_GPA))+geom_point()+geom_smooth(method="lm")

p3<-ggplot(gpa, aes(verb_SAT, univ_GPA))+geom_point()+geom_smooth(method="lm")


grid.arrange(p1, p2, p3, nrow=2)

```


c) Describe any relationship on your scatterplot: its direction, its strength and its shape. Justify your description briefly.

All three of theme exhibit a positive correlation, because as the x values increase the university GPA also increase. For high GPA against university GPA the relationship appears at least moderate, perhaps strong, but for verbal SAT and math SAT, it appears moderate as the points don't necesssarily trend along the line of best fit.

d) Fit a linear regression for predicting university GPA from high-school GPA and display the results.

```{r}
model1<-lm(univ_GPA~high_GPA, gpa)
summary(model1)
```
The model is given by: 

$$\hat{y} = 1.0968 + 0.6748 \times  high GPA$$



where $\hat{y}$ = university GPA. Note that $\hat{y}$ is an unbiased estimator of $E(y_i)$

For every unit increase in high school GPA there is an increase in the mean change of college GPA by a factor of 0.6748.

d) Two students have been admitted to university. One has a high school GPA of 3.0 and the other a high school GPA of
3.5. Obtain suitable intervals that summarize the GPAs that each of these two students might obtain in university.

You are being asked to find a confidence interval based on their scores.

```{r}
newdata=data.frame(high_GPA =c(3.0,3.5))
 predict(model1, newdata, interval="confidence",level=0.95) 
```

If not given an $\alpha$ assume that it is 0.05. This is why I have a 95% confidence interval; it is $(1-\alpha)$% interval. 

* Now obtain a regression predicting university GPA from high-school GPA as well as the two SAT scores. Display your results.

```{r}
model2<-lm(univ_GPA~high_GPA + verb_SAT + math_SAT, gpa)
summary(model2)
```



Just for curiousity lets look at the pairs plot:

```{r}
pairs(~high_GPA+verb_SAT+math_SAT)
```

e) Test whether adding the two SAT scores has improved the prediction of university GPA. What do you conclude?

We might say that the two SAT scores has not improved the model:

1) It shrinks the effect of the significant covariate (high_GPA)
2) Good answer: If you perform a hypothesis test on both (i.e. verbal SAT and math SAT) you will see that none of them are actually significant.
3) The F statistic shrunk, so the global significant is also lowered by adding two insignificant covariates (i.e. 159.6 to 55.77)


## Categorical variables in multiple regression:

Let's look at mtcars:

```{r}
mtcars
```


Suppose we want to examine the association between a continuous response variable (mpg) and a categorical predictor variable (cyl). We learned that a scatter plot is plausible course of action to display the relationship in bivariate statistics.

```{r}
#if you use a scatterplot it looks odd
ggplot(mtcars, aes(x=as.factor(cyl), y=mpg))+geom_point()

#we will use a boxplot 
ggplot(mtcars, aes(x=as.factor(cyl), y=mpg))+geom_boxplot()
```

Perhaps I think that the mpg is related to how heavy a car is. This may result in my using cyl and wt. 


```{r}
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point()+geom_smooth(method="lm")
```


```{r}
model2<-lm(mpg~wt + as.factor(cyl), dat=mtcars)
summary(model2)
```
Based on this we can see that weight and cylinder class are both significant. The p-values for both are less than 0.05. However, I will check to see how wt is related to cyl. 

```{r}
ggplot(mtcars, aes(x=as.factor(cyl), y=wt))+geom_boxplot()
```

Note: When you have two variables that are potentially associated but one is categorical, you may think to perform a chi-squared test for independence. 

What is the interpretation of a categorical coefficient?

Here we an see that the reference group is "4" within the cylinders. Why? (We have a decrease in the mpg when you move to 6 and 8 cylinder classes from class 4). It is the mean change you get when you move from class 4 to 6 or class 4 to 8. 

Interpretation of the coefficient must be made in comparison to the reference group (4 cylinders), which is absorbed in the intercept coefficient. 

How do we determine significance?

$$H_0: \text{All}  \hspace{0.05in} \beta_i= 0$$
$$H_1: \text{At least one $\beta_i$ is not 0}$$

This is the same as it always was. Here you have to check the coefficients for level 6 and 8. Not all levels are significant, so this is why we perform the test separately. 
