---
title: "Intro to R Markdown and Tidyverse"
author: "Nnenna Asidianya"
date: "March 14 2022"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```



Read in mortality rates for Canada. These data come from the [Human Mortality Database](https://www.mortality.org/), which is a great source for studying mortality in high-income countries. 

```{r}
library(tidyverse)
dm <- read_table("CAN_Mx_1x1.txt", skip = 0)
head(dm)
attach(dm)
```


## Important tidyverse functions

You should feel comfortable using the following functions
- `pivot_longer` and `pivot_wider`


#Change a string to numeric

```{r}
dm <- 
  dm %>% 
  mutate(Age = as.numeric(Age), 
         Female = as.numeric(Female), 
         Male = as.numeric(Male),
         Total = as.numeric(Total))
```

## Pivoting

We often need to switch between wide and long data format. The `dm` tibble is currently in wide format. To get it in long format we can use `pivot_longer`

```{r}
dm_long<-dm %>% pivot_longer(cols = Female:Total, names_to = "sex", values_to = "rate" )

dm_wide <-dm_long%>% pivot_wider(names_from = "sex", values_from="rate")
```



## Using ggplot

You can plot things in R using the base `plot` function, but plots using `ggplot` are much prettier. 

Say we wanted to plot the mortality rates for 30 year old males over time. In the function `ggplot`, we need to specify our data (in this case, a filtered version of dm), an x axis (Year) and y axis (Male). The axes are defined withing the `aes()` function, which stands for 'aesthetics'.

First let's get our data:

```{r}
dm_to_plot <- dm %>% 
  filter(Age==30) 

dm_to_plot <- dm_to_plot %>% 
  pivot_longer(Female:Total, "sex", values_to = "mx") %>% 
  filter(sex!="Total")
```

Now start the ggplot:

```{r}
ggplot(dm_to_plot, aes(Year, mx, color = Age, group = Age)) + 
  geom_line()+
  ylab("Mortality rate")+
  facet_grid(~sex)+
  ggtitle("Mortality rates for 30 Canadians, 1921-2011")+      
  scale_color_gradient(low = "blue", high = "red")+ theme_classic()
```

Let's see what happens if we filter based on Male

```{r}
dm_plot2<-dm_to_plot %>% filter(sex=="Male")

ggplot(dm_plot2, aes(Year, mx, color = Age, group = Age)) + 
  geom_line()+
  ylab("Mortality rate")+
  ggtitle("Mortality rates for 30 Canadians, 1921-2011")+      
  scale_color_gradient(low = "blue", high = "red")+ theme_classic()
```
